{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-16 thesis/education.tex
bias and pebbles retro

@@

%%%% pintos %%%% (?? this is psu not pintos)
% survey questions i WISH i had asked
% - did you have any technical difficulties w landslide that i had to intervene on
% mb anything else from timmys 2nd latest email

@@

>>>>

During the last week of P2 at CMU, I held several ``Landslide clinic'' sessions
(basically office hours, but given a different name to remind students
to limit themselves to questions a normal TA couldn't answer),
where students could receive in-person technical and/or moral support.
Collecting study information during these sessions was not included in the IRB protocol.
In the PSU study, I had returned to Pittsburgh shortly after giving the lecture,
so technical support for PSU students was limited to email correspondence.

@@

As long as an educational user study is run on a volunteer basis,
one cannot completely avoid selection bias:
those with enough free time to participate are more likely to be the most capable students already,
who are least in need of the extra debugging help.
This was especially evident in the annotation-required P3 study from \cite{landslide},
in which only 5 groups volunteered (15\% of the 34 total who submitted P3 that semester),
among which 2 had enough after just the annotation phase and did not continue to do any in-depth testing.
In contrast, since switching to the automatically-annotatable P2,
the participation rate rose to NN\% % TODO
(XX out of YY P2-submitting groups across ZZ semesters (TODO actually count them but obviously it will be higher)). % TODO
This alone could be seen as a major step in mitigating said selection bias.

Beyond participation, I anticipated several forms of bias affecting the survey responses.
First of all, I suspected that students might feel compelled to be overly polite in their answers
(whether consciously or subconsciously).
I attempted to counteract this by concluding my survey link emails with,
{\em Please try to answer honestly rather than flatteringly--if any part
of the experience was bad for you, I want to hear about it to make
Landslide better in the future!}
It's also possible that survey respondents % respondence
were more likely to be those who enjoyed Landslide the most,
meaning I might not hear as much negative feedback as I should.
Table~\ref{tab:TODO} % TODO -- make a table or graph showing correlation between bugfinding and who responded on the survey
is my attempt to dispel this possibility.
% TODO
% possible flaws in table: students who found bugs might have had either a good (bug!) or bad (false positive) experience
% could studence who didn't find bugs also have had a good experience? (think of how...)

% TODO: is this ok to report without IRB approval? ask dave
I cannot think of any reason gender, racial, or other demographic-based bias might affect this study.
In fact, according to my memory alone, the proportion of women among students
attending the Landslide clinics (\sect{\ref{sec:education-pebbles-recruiting}})
was considerably higher than the class's overall gender ratio,
although I did not record any such statistics scientifically.

@@

The survey was introduced into the study only in time for two semesters' worth of student responses,
after several iterations of collecting only usage and bug report snapshots.
Apart from the obvious improvement of having been surveying students from the beginning,
the following questions could have improved the survey.
\begin{enumerate}
	\item {\em Did you have any technical difficulties with Landslide that I had to intervene on,
		whether in person or over email?}
		(Some students reported this in the ``false positives'' question,
		although fewer than I helped overall, so others must have not mentioned it.)
		Comparing answers to this question across subsequent semesters would give a sense
		of how much Landslide's stability was improving over time
		and whether it was mature enough for unsupervised use in the future.
	\item {\em How well do you feel you understood the research challenges explained in the lecture?}
		and,
		{\em How well do you feel a user should need to understand same in order to benefit from Landslide's bug reports?}
		(Answers on a scale from ``Not at all; Landslide is a magic black box to me''
		to ``I'm ready to work on research in this field myself''.)
		These questions would help fine-tune the lecture material and user guide to maximize student comfort,
		and potentially also corroborate the claim that Landslide is accessible even to novice users.
	\item {\em What additional debugging information would you want displayed on the preemption traces?}
		Knowing now that interpreting preemption traces was a sticking point for many users,
		I would hope to identify the most wished-for features to know what to prioritize improving.
		This could also assess students' understanding of what kinds of information
		would or would not be reasonable for Landslide to record and report.
\end{enumerate}

Some students (between 0 and 2 per semester)
emailed me during P3
to ask if they could test their kernels with Landslide just like their thread libraries.
I answered each by explaining that it would take more effort on their part,
and if they were still interested,
guided them through the annotation process on a case-by-case basis.
This process was not included in the IRB-approved study protocol,
so I collected no results from them.
If I had planned in advance, I could have supported this ``bonus stage'' officially,
and further surveyed the brave volunteers about how P3 Landslide could be made more generally accessible.

Finally, to evaluate whether the experience of using Landslide
left the students with any lasting lessons learned,
a follow-up survey could have been given one or two years later.
Such a survey would ask, for example,
{\em Have you encountered any debugging problems since finishing OS that made you wish for a tool like Landslide?}
and
{\em Do you feel the way you think about testing, debugging, and program correctness
has been influenced in any way by using Landslide?}.
